### 2020-11-14 Thesis-Chapter5-ThreatModel.tex
Finish risk analysis

\modelitem{Foreign intelligence}{Foreign intelligence agencies and \acp{APT} are always looking for means to escalate
access and tamper with and steal information. They have world-class expertise and computing power, which they will apply
to both individual devices and backend systems.}

@@

% What are we building?: proposal DFDs
% What can go wrong?: elicitation of threats

\section{Basic Data at Rest}

@@

\threattablebegin{Spoofing}{spoofing}{Spoof}
  S1 & Device is spoofed to LE via plant
  & Attackers plant a device to get LE to unlock it for them.
  & Require the device to be in the forensics department itself the entire time it is unlocked.
  \\ \hline
  S2 & Device is spoofed to LE via forged DID
  & Attackers rig a device to report a different DID and record the entered unlock artifacts, allowing them into a
    target device.
  & Add a hardware integrity check or alternatively unlock the device in a Faraday cage (to prevent telemetry), extract
    its data, and destroy it (to prevent recovering stored data).
  \\ \hline
  S3 & LEO is spoofed to LEA
  & Attackers pose as a LEO while the LEA relays back unlock artifacts.
  & Require the device to be in the forensics department itself the entire time it is unlocked.
  \\ \hline
  S4 & LEA is spoofed to vendor during key exchange
  & Attackers pose as an LEA in order to spoof future access requests.
  & Vendor key exchange process involves thorough in-person verification.
  \\ \hline
  S5 & LEA is spoofed to vendor during access request
  & Attackers pose as a LEA and submit an access request. Requires forged legal documents and a stolen LEA private key
    (unless attacker is an insider or has executed a spoofed key exchange).
  & Vendor independently verifies all legal documents. Vendor supports a key revocation and update process that involves
    the same level of in-person verification.
  \\ \hline
  S6 & Manufacturer is spoofed to vendor
  & Attackers pose as the manufacturer to inject malicious data into the DID-DSK database. Must be combined with other
    attacks to be of use.
  & Vendor authenticates data transfers between manufacturing and corporate (does not address a compromised manufacturer
    network or an inside threat).
\threattableend

@@

\threattablebegin{Tampering}{tampering}{Tamper}
  T1 & Defensive user corrupts DSK, DID, or token
  & LE attempts at perform \ac{EA} will fail due to corrupt data. Requires compromise of the Secure Enclave or advanced
    hardware manipulation tools.
  & Since these values are known at manufacturing time and never change, they can be made read-only. (The symmetric key
    variant of \ldawmsr requires updating the symmetric equivalent of the public DSK, and is therefore more vulnerable
    to this threat.)
  \\ \hline
  T2 & Defensive user destroys the \ac{EA} hardware interface
  & LE attempts to perform \ac{EA} will fail due to damaged hardware.
  & Manufacture the device to require destructive disassembly to access the interface. This increases the cost
    associated with deploying and using the system.
  \\ \hline
  T3 & LE insider tampers with DID before digital signature
  & LEA receives an unauthorized device's unlock artifacts. Can be combined with other attacks to compromise a device.
  & Randomize DID's to make them unguessable, meaning a target device must be in possession in order for the attacker to
    know what DID to use. Input to the vendor's \ac{HSM} could also include data from the legal documents, such as
    device S/N; the \ac{HSM} would hash the value and verify a match before returning the unlock artifacts.
  \\ \hline
  T4 & Attacker tampers with hardware transparency mechanism
  & The device user is unaware that the device was unlocked. Highly implementation dependent, but certainly requires
    advanced hardware manipulation tools.
  & Manufacture the device to require destructive disassembly to access the interface.
\threattableend

@@

\threattablebegin{Repudiation}{repudiation}{Repudiation}
  R1 & Attacker or authoritarian government repudiates unlock attempt
  & The user is unaware that someone tried to access their device.
  & Manufacture the device to require destructive disassembly to access the interface, or use the same hardware
    transparency mechanism to mark mere access attempts as well.
  \\ \hline
  R2 & Authoritarian government repudiates demanding vendor assistance
  & Governments not respecting the legal framework governing the system can bully the vendor and try to force access
    while maintaining deniability.
  & This is close to being an out of scope threat under the ``rule of law'' exception. However, one could design the
    system to require requests to be recorded on a public ledger. This opens up several new difficulties.
  \\ \hline
  R3 & LEA repudiates having unlocked a device
  & The device is undeniably unlocked, but the LEA claims it was not them, allowing possible violations of civil
    liberties and harming transparency.
  & The escrow agent could require some identifying information before unlocking, such as the LEA's public key digitally
    signed by the vendor.
  \\ \hline
  R4 & Vendor repudiates having responded to an access request
  & The device is undeniably unlocked, but the vendor claims they did not provide the unlock artifacts, allowing the
    vendor to cover up leaks or unsavory policy decisions by suggesting the device was instead hacked. Harms
    transparency.
  & The same identifying information used in \ii{R3} would implicate the vendor (assuming the integrity of the vendor's
    signing key).
\threattableend

@@

\threattablebegin{Information Disclosure}{disclosure}{Information \\ Disclosed}
  I1 & Attacker exposes the DSK, DID, the encrypted class key, or hashed token
  & By design, none of the information stored on the phone is sensitive. The most consequential information that could
    be stolen is the DID, which could be paired with other attacks. Disclosing the DID requires either disassembly and
    time vaulting or advanced hardware inspection tooling.
  & Sensitivity of this data has already been mitigated as much as possible.
  \\ \hline
  I2 & Attacker intercepts unlock artifacts between LEA \ac{HSM} and device
  & The attacker is able to unlock the target device alone. Must be combined with a spoofed DID to be of use to other
    devices (see \ii{S2}). The could record the values to replay them after the phone has been returned.
  & See \ii{S2}. If choosing not to destroy the device, to prevent replay attacks, a new token could be generated and
    stored in the vendor HSM, with the new hash installed during the unlock process.
  \\ \hline
  I3 & Attacker steals the LEA private key
  & The attacker is now able to spoof the LEA and decrypt returned unlock artifacts. Without the devices (which are
    already in LE possession), the artifacts are not useful. See \ii{S4} and \ii{S5} for LEA spoofing attacks.
  & The key is stored in an \ac{HSM}, which has extensive security measures. See \ii{S5} for a revocation and update
    mitigation.
  \\ \hline
  I4 & Attacker steals the DID-DSK database
  & This is the big prize, the largest source of centralized risk. With the database, attackers could directly
    compromise (or more likely, sell access to) any mobile device using this system in their possession, after
    disassembly, and after the time vault.
  & \ii{R3} defines a mitigation that requires use of a vendor private key to create a digital signature as part of the
    authentication process. If used, that key would have to be compromised as well for the database to be of use.
    However, if the database is compromised, that key probably is too. Savage argues that \ac{HSM}s, though imperfect,
    actually do provide very strong assurance. Responses could be rate-limited, and leaking an entire database once it's
    in the \ac{HSM} could not go unnoticed. Getting the data in may be harder---see \ii{I5}.
  \\ \hline
  I5 & Attacker steals device unlock artifacts between manufacturing and vendor \ac{HSM}
  & Intercepting the unlock artifacts at the source achieves the same effect as \ii{I4}, only the data set is not
    complete. Instead of granting the ability to unlock any device, including a specific target, this is suited for
    getting access down the line or for sale.
  & The unlock artifacts generation process must be treated as sensitively as the \ac{HSM} itself. The artifacts should
    be encrypted immediately and periodically transported to the corporate \ac{HSM} on physical media. This threat is
    difficult to definitively mitigate.
\threattableend

@@

\threattablebegin{Elevation of Privilege}{privilege}{Privilege Gained}
  E1 & Foreign intelligence or organized crime recruits an insider
  & The primary attacker gains inside access through coercion or bribery, making any other attack easier.
  & Limit the employees, accounts, and computers involved in any element of the system. Perform background checks on
    each employee and maintain detailed logs.
  \\ \hline
  E2 & Unexposed hardware interface allows access to Secure Enclave
  & The attacker gains access to Secure Enclave operations and data, undermining the device and achieving unlock without
    following the \ac{EA} protocol.
  & The time vaulting protocol and all communication over the interface must be thoroughly tested, and if possible,
    formally proven to be logically sound. Analysis must include side channel attacks to affect the time vaulting or
    leak internal data.
\threattableend

@@

How much risk does \ldawmsr introduce? The answer varies based on how many mitigations are implemented. If the device is
required to be unlocked on LEA premises and destructive disassembly is chosen, several threats are significantly
reduced. The irreducible risks are those posed by the DID-DSK database and the Secure Enclave itself.

\section{Response to Frameworks}

% Or just, relationship to my goals?

% How does it answer (or not) committee_decrypting_2018's framework questions? What needs improvement?

% Reflection on how much progress this represents

>>>>

\modelitem{Foreign intelligence}{Foreign intelligence agencies and \acp{APT} are always looking for means to escalate
access and tamper with and steal information. They have world-class expertise and computing power, which they will apply
to both individual devices and backend systems. (Here I am referring to \ii{foreign} foreign intelligence agencies, as
opposed to \ii{domestic} foreign intelligence agencies such as the \acs{CIA} or \acs{NSA} in the U.S., which are assumed
to operate within---though push the boundaries of---the legal framework.)}

@@

\section{Basic Data at Rest}
\label{sec-basic-dar}

@@

\threattablebegin{Spoofing}{spoofing}{Spoof}
  S1 & Device is spoofed to LE via plant
  & Attackers plant a device to get LE to unlock it for them.
  & Require the device to be in the forensics department itself the entire time it is unlocked.
  \\ \hline
  S2 & Device is spoofed to LE via forged DID
  & Attackers rig a device to report a different DID and record the entered unlock artifacts, allowing them into a
    target device.
  & Add a hardware integrity check or alternatively unlock the device in a Faraday cage (to prevent telemetry), extract
    its data, and destroy it (to prevent recovering stored data).
  \\ \hline
  S3 & LEO is spoofed to LEA
  & Attackers pose as a LEO while the LEA relays back unlock artifacts.
  & Require the device to be in the forensics department itself the entire time it is unlocked.
  \\ \hline
  S4 & LEA is spoofed to vendor during key exchange
  & Attackers pose as an LEA in order to spoof future access requests.
  & Vendor key exchange process involves thorough in-person verification. Assume only a handful of digital forensics
    laboratories nationwide are authorized to participate.
  \\ \hline
  S5 & LEA is spoofed to vendor during access request
  & Attackers pose as a LEA and submit an access request. Requires forged legal documents and a stolen LEA private key
    (unless attacker is an insider or has executed a spoofed key exchange).
  & Vendor independently verifies all legal documents. Vendor supports a key revocation and update process that involves
    the same level of in-person verification. Assume only a handful of digital forensics laboratories nationwide are
    authorized to participate.
  \\ \hline
  S6 & Manufacturer is spoofed to vendor
  & Attackers pose as the manufacturer to inject malicious data into the DID-DSK database. Must be combined with other
    attacks to be of use.
  & Vendor authenticates data transfers between manufacturing and corporate (does not address a compromised manufacturer
    network or an inside threat).
\threattableend

@@

\threattablebegin{Tampering}{tampering}{Tamper}
  T1 & Defensive user corrupts DSK, DID, or token
  & LE attempts at perform \ac{EA} will fail due to corrupt data. Requires compromise of the Secure Enclave or advanced
    hardware manipulation tools.
  & Since these values are known at manufacturing time and never change, they can be made read-only. (The symmetric key
    variant of \ldawmsr requires updating the symmetric equivalent of the public DSK, and is therefore more vulnerable
    to this threat.)
  \\ \hline
  T2 & Defensive user destroys the \ac{EA} hardware interface
  & LE attempts to perform \ac{EA} will fail due to damaged hardware.
  & Manufacture the device to require destructive disassembly to access the interface. This increases the cost
    associated with deploying and using the system.
  \\ \hline
  T3 & LE insider tampers with DID before digital signature
  & LEA receives an unauthorized device's unlock artifacts. Can be combined with other attacks to compromise a device.
  & Randomize DIDs to make them unguessable, meaning a target device must be in possession in order for the attacker to
    know what DID to use. Input to the vendor's \ac{HSM} could also include data from the legal documents, such as
    device S/N; the \ac{HSM} would append the S/N to the DID and hash the result, verifying a match with the same stored
    hash before returning the unlock artifacts.
  \\ \hline
  T4 & Attacker tampers with hardware transparency mechanism
  & The device user is unaware that the device was unlocked. Highly implementation dependent, but certainly requires
    advanced hardware manipulation tools.
  & Manufacture the device to require destructive disassembly to access the interface.
\threattableend

@@

\threattablebegin{Repudiation}{repudiation}{Repudiation}
  R1 & Attacker or authoritarian government repudiates unlock attempt
  & The user is unaware that someone tried to access their device.
  & Manufacture the device to require destructive disassembly to access the interface, or use the same hardware
    transparency mechanism to mark mere access attempts as well.
  \\ \hline
  R2 & Authoritarian government repudiates demanding vendor assistance
  & Governments not respecting the legal framework governing the system can bully the vendor and try to force access
    while maintaining deniability.
  & This is close to being an out of scope threat under the ``rule of law'' exception. However, one could design the
    system to require requests and responses to be recorded on a public ledger. This creates several new difficulties.
  \\ \hline
  R3 & LEA repudiates having unlocked a device
  & The device is undeniably unlocked, but the LEA claims it was not them, allowing possible violations of civil
    liberties and harming transparency.
  & At a high level, the system should require a long paper trail before access can be granted. At a low level, the
    escrow agent could require and store some identifying information before unlocking, such as the LEA's public key
    digitally signed by the vendor, which provides non-repudiation assuming independent access to the device. See also
    the mitigation in \ii{R2}.
  \\ \hline
  R4 & Vendor repudiates having responded to an access request
  & The device is undeniably unlocked, but the vendor claims they did not provide the unlock artifacts, allowing the
    vendor to cover up leaks or distasteful policy decisions by suggesting the device was instead hacked. Harms
    transparency.
  & Assuming independent access to the device, the hardware transparency mechanism should confirm whether the \ac{EA}
    process was used, and the same identifying information used in \ii{R3} would implicate the vendor. See also the
    mitigation in \ii{R2}.
\threattableend

@@

\threattablebegin{Information Disclosure}{disclosure}{Information \\ Disclosed}
  I1 & Attacker exposes the DSK, DID, the encrypted class key, or hashed token
  & By design, none of the information stored on the phone is sensitive. The most consequential information that could
    be stolen is the DID, which could be paired with other attacks. Disclosing the DID requires either disassembly and
    time vaulting or advanced hardware inspection tooling.
  & Sensitivity of this data has already been mitigated as much as possible.
  \\ \hline
  I2 & Attacker intercepts unlock artifacts between LEA \ac{HSM} and device
  & The attacker is able to unlock the target device alone. Must be combined with a spoofed DID to be of use to other
    devices (see \ii{S2}). The attacker could record the values to replay them after the phone has been returned.
  & See \ii{S2}. If choosing not to destroy the device, to prevent replay attacks, a new token could be generated and
    stored in the vendor HSM, with the new hash installed during the unlock process.
  \\ \hline
  I3 & Attacker steals the LEA private key
  & The attacker is now able to spoof the LEA and decrypt returned unlock artifacts. Without the devices (which are
    already in LE possession), the artifacts are not useful. See \ii{S4} and \ii{S5} for LEA spoofing attacks.
  & The key is stored in an \ac{HSM}, which has extensive security measures. See \ii{S5} for a revocation and update
    mitigation.
  \\ \hline
  I4 & Attacker steals the DID-DSK database
  & This is the big prize, the largest source of centralized risk. With the database, attackers could directly
    compromise any mobile device using this system in their possession, after disassembly, and after the time vault.
    They may also try to sell access to the data.
  & \ii{R3} defines a mitigation that requires use of a vendor private key to create a digital signature as part of the
    authentication process. If used, that key would have to be compromised as well for the database to be of use.
    However, if the database is compromised, that key probably is too. Savage argues that \acp{HSM}, though imperfect,
    actually do provide very strong assurance. Responses could be rate-limited, and leaking an entire database once it's
    in the \ac{HSM} could not go unnoticed. Getting the data in may be harder---see \ii{I5}.
  \\ \hline
  I5 & Attacker steals device unlock artifacts between manufacturing and vendor \ac{HSM}
  & Intercepting the unlock artifacts at the source achieves the same effect as \ii{I4}, only the data set is not
    complete. Instead of granting the ability to unlock any device, including a specific target, this is suited for
    getting access down the line or for sale.
  & The unlock artifacts generation process must be treated as sensitively as the \ac{HSM} itself. The artifacts should
    be encrypted immediately and periodically transported to the corporate \ac{HSM} on physical media. This threat is
    difficult to definitively mitigate.
  \\ \hline
  I6 & Attacker steals device \ac{plaintext} from LEA after unlock process
  & Assuming the attacker doesn't care that the LEA gets access to plaintext, they could ``help'' the LE investigation
    in order to access the data later either when it comes out as public evidence or through compromising the LEA
    itself. This may be useful to authoritarian regimes, foreign intelligence, and organized crime.
  & LEAs must have strong information security of their own. Disclosure of evidence is a natural side effect of
    prosecution, so LEAs must practice discretion in choosing which devices to unlock in case mere information discovery
    poses a great risk in itself.
\threattableend

@@

\threattablebegin{Elevation of Privilege}{privilege}{Privilege Gained}
  E1 & Foreign intelligence or organized criminals recruit an insider
  & The primary attacker gains inside access through coercion or bribery, making any other attack easier.
  & Limit the employees, accounts, and computers involved in any element of the system. Perform background checks on
    each employee and maintain detailed logs.
  \\ \hline
  E2 & Unexposed hardware interface allows access to Secure Enclave
  & The attacker gains access to Secure Enclave operations and data, undermining the device and achieving unlock without
    following the \ac{EA} protocol.
  & The time vaulting protocol and all communication over the interface must be thoroughly tested, and if possible,
    formally proven to be logically sound. Analysis must include side channel attacks to affect the time vaulting or
    leak internal data.
  \\ \hline
  E3 & An \ac{HSM} vulnerability allows access to protected data
  & This is a technical route to achieving \ii{I4}, disclosure of the DID-DSK database.
  & The \ac{HSM} should be on an air-gapped network disconnected to the internet, physically guarded, and regularly
    patched. There is no defense against an unknown vulnerability, so security ultimately comes down to incident
    detection and response.
\threattableend

@@

How much risk does \ldawmsr introduce? The answer lies in a range depending on how many mitigations are implemented.
Some threats introduce inherently low risk due to their difficult nature and low pay off. Others are higher risk but can
be reasonably mitigated. Some threats specifically target transparency, and are difficult to address without trust.
Finally, there are some risks that are irreducible. Each of classes are discussed in turn.

\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, and \ii{E1} \ii{E2} are each low risk either because they
fail to produce useful results, they can be well mitigated, or they are a threat that is already present in
state-of-the-art designs as depicted in \mysec{sec-basic-dar}. It is crucial for a design to consider and mitigate the
risk posed by these threats as part of the design process, but they do not represent fatal flaws to a design. Except in
combination with other threats, they are not worthy more concern here.

\ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are each higher risk threats to the
security, trustworthiness, and effectiveness of the proposal, but can be addressed through a set of related mitigations:
require destructive disassembly of the device to access the \ac{EA} hardware interface, require the device to be
physically present on LEA premises the entire duration it is unlocked, and grant only a handful of LE facilities the
lawful right to use the \ac{EA} mechanism. Requiring destructive disassembly is a technical measure that entirely
precludes certain attacks; requiring the device to be on LEA premises is a procedural measure that prevents attackers
from manipulating LE into achieving their own ends; limiting the number of LE facilities that can participate prevents
LE impersonation attacks. Together, they make the proposal considerably harder to attack.

\ii{R2}, \ii{R3}, and \ii{R4} are repudiation attacks that belie the system's potential for abuse. The system relies on
legal process and cooperation of several parties. But what if the legal process involves secret courts, secret
investigations, and secrecy orders on the vendor \cite{shamsi_2011}? Destructive disassembly or the weaker hardware
transparency mechanism provide non-repudiation that the device was unlocked, if and only if one has physical access to
the device. Technically enforced transparency into device access requests and responses would grant non-repudiation at a
broader level. \ldawmsr does not include a mechanism for the feature, though other \ac{DAR} \ac{EA} proposals do
\cite{phan_key_2017} \cite{servan_schreiber_jje_2020}. Unfortunately, building this feature into a proposal makes it
potentially vulnerable the hawkish lawmaker, who may consider it an anti-feature.

Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, each representing in slightly different ways the
disclosure of unlock artifact data, the DID-DSK database. This is the truly difficult challenge---centralizing so much
sensitive data makes it an extremely high value target for foreign intelligence and cyber criminals. The vendor is in a
position where they need to protect this database from the most advanced attackers in the world while regularly updating
and accessing it. \Acp{HSM} and strong security policies would make that a difficult task, but not impossible, and one
has to assume that the data would leak eventually. What would leaked unlock artifacts mean? It depends who steals it.
Criminals would use it for personal and financial gain; foreign intelligence and authoritarian regimes would use it to
spy on their enemies. In the worst-case scenario, anyone with some specialized tools and a willingness to break the law
could get into any device under the scheme, provided they destructively disassemble it and wait through the time vault.

It is important to give \ldawmsr (and similar \ac{DAR} \ac{EA} proposals) its due. The worst case scenario still
precludes mass surveillance and includes non-repudiation that access occurred. This is a significant gain for \ac{EA}
proposals in terms of transparency and protection of civil liberties.

\section{Achievement of Goals}

% Consider goals, threat actors, and loop-back to out of scope items, because they aaaaactually do matter.

% Economics: vastly increases costs and barrier to entry for vendors.
